The truncated KdV framework for modeling anomalous waves induced by abrupt depth changes

### 2019-08-07 paper2_JNLS/paper.tex
NM

Rogue waves are abnormally large surface waves, defined by oceanographers as those that exceed twice the significant wave height. Though such waves were once dismissed as myth, they have now been recorded in oceans across the globe and pose a recognized threat to seagoing vessels and naval structures. Rogue waves, also variously known as freak or anomalous waves, have been observed both in deep and shallow water, and can be triggered by a variety of mechanisms, including anomalous wind-forcing, focusing due to variable bathymetry, and a deep-water modulational instability that goes by the name Benjamin-Feir. The common tie between these mechanisms is their ability to generate non-normal statistics in the surface displacement; i.e.~When governed by Gaussian statistics, the likelihood of a rogue wave is extremely low, but these rare events can become much more common when statistics deviate from Gaussian.

A recent series of laboratory \cite{bolles2019anomalous} and numerical investigations \cite{viotti2014} have demonstrated the emergence of anomalous wave statistics from variations in bottom topography.

% Since studies are 1D topography, they offer bare minimum set of conditions able to generate anomalous waves; offers paradigm problem that can be studied scrutinized; shares features with more complex situations.

% ideal, paradigm, canonical

>>>>

Rogue waves are abnormally large surface waves, defined by oceanographers as those that exceed twice the significant wave height. Though such waves were once dismissed as myth, they have now been recorded in oceans across the globe and pose a recognized threat to seagoing vessels and naval structures. Rogue waves, also variously known as freak or anomalous waves, have been observed both in deep and shallow water, and can be triggered by a variety of mechanisms, including anomalous wind-forcing, focusing due to variable bathymetry, and a deep-water modulational instability that goes by the name Benjamin-Feir. The common tie between these mechanisms is their ability to generate non-normal statistics in the surface displacement; i.e.~When governed by Gaussian statistics, the likelihood of a rogue wave is extremely low, but these rare events can become much more common when surface statistics deviate from Gaussian.

A recent series of laboratory \cite{bolles2019anomalous} and numerical investigations \cite{viotti2014} have demonstrated the emergence of anomalous wave statistics from variations in bottom topography. Since topographical variations are strictly one-dimensional in these studies, they can be considered as offering a bare minimum set of conditions capable of generating anomalous wave statistics. In particular, the more complex mechanism of focusing by 2D bottom topography absent. The studies thus offer the simplest conceivable system that exhibits the anomalous statistical behavior present in more complex systems, like the ocean, but in a tractable system that could be analyzed and quantified.
% ideal, paradigm, canonical, tractable
